---
title: 'Metabolic Model Update: Base Model Totals'
author: 'Matthew Onorato'
version: '0.0.1'
date: '10/23/2020'
maintainer: 'Matthew Onorato'
email: 'monorato2@gmail.com' > 'monorato4398@sdsu.edu'
status: 'final'
output: html_document
---

DESCRIPTION: This script creates a stacked bar plot with new model counts   
             stacked on top of old model counts. More aesthetically pleasing
             than the Matplotlib bar plot.

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(gridExtra)
library(grid)
library(gridtext)

rm(list = ls())

work_dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(work_dir)
getwd()

knitr::opts_chunk$set(echo = TRUE)
```


Stacked barplot
-values were inserted manually (no COBRA library in R for metabolic modeling) 
```{r}
totals <- tribble(~'Model', ~'Reactions', ~'Metabolites', ~'Genes',
                  'iMtb_H37Rv_1132', 181, 80, 124,
                  'iEK1008', 1226, 998, 1008) %>% pivot_longer(Reactions:Genes)

totals$name <- factor(totals$name, levels=c('Reactions', 'Metabolites', 'Genes'))
totals$Model <- factor(totals$Model, levels=c('iMtb_H37Rv_1132', 'iEK1008'))

rmg <- ggplot(totals, aes(x=name, y=value, fill=Model)) + 
       geom_bar(position='stack', stat='identity', width=0.5, color='white') +
       theme_classic() + labs(x='', y='Model Total') +
       scale_y_continuous(breaks=c(0, 200, 400, 600, 800, 1000, 1200, 1400),
                          expand=c(0, 0)) +
       theme(legend.position=c(0.825, 0.93), legend.title=element_blank(),
             legend.background=element_blank(),
             legend.text=element_text(size=10, face='bold'),
             legend.box.background=element_rect(color='grey'),
             legend.spacing.y=unit(0, 'mm'),
             axis.text.x=element_text(color='black', size=10, face='bold',
                                      margin=margin(5, 0, -18, 0)),
             axis.text.y=element_text(color='black', size=10, face='bold'),
             axis.title.y=element_text(size=10, face='bold')) +
       scale_fill_manual(values=c('#4AA5C2', '#A3D9E6'))
```


Accompanying table (values also inserted manually)
```{r}
table_df <- tribble(~'Reactions', ~'Metabolites', ~'Genes',
                    '1407 (+14.8%)', '1078 (+8.0%)', '1132 (+12.3%)',
                    '+211', '+112', '+133',
                    '-30', '-32', '-9',
                    '1226', '998', '1008')

row_colors <- matrix('white', nrow(table_df), ncol(table_df))
row_colors[1, 1:ncol(row_colors)] <- '#4AA5C2'
row_colors[4, 1:ncol(row_colors)] <- '#A3D9E6'

tbl_thm <- ttheme_minimal(
  core=list(bg_params=list(fill=row_colors, alpha=0.8),
            fg_params=list(fontface=3L, fontsize=10)),
  colhead=list(fg_params=list(fontface=4L, fontsize=10)),
  padding=unit(c(22, 4), 'mm'))

table <- tableGrob(table_df, theme=tbl_thm, rows=NULL, cols=NULL)
```


Combines barplot, table, and caption into a single figure
```{r}
rmg_tb_cap <- grid.arrange(rmg, table, heights=c(1.1, 0.3), widths=c(0.2, 2),
                           layout_matrix=rbind(c(1, 1), c(NA, 2)))

ggsave('1_base_model_totals.jpg',
       plot=rmg_tb_cap, height=5.5, width=6, unit ='in')

# cap <- textbox_grob('**Figure 1. Net Increase in Reactions, Metabolites, and
#                     Genes after Updating iEK1008 to iMtb_H37Rv_1132.** Overall,
#                     the curation process led to a net gain of 181 reactions, 80
#                     metabolites, and 124 genes representing increases of 14.8%,
#                     8.0%, and 12.3%, respectively. 32 disconnected metabolites
#                     and 9 disconnected genes were removed and 30 reactions were
#                     removed due to duplicity, lack of experimental evidence, or
#                     Biolog result contradiction.',
#          gp=gpar(fontsize=10), x=unit(0.55, 'npc'), y=unit(0.5, 'npc'),
#          width=unit(0.9, 'npc'))
# 
# rmg_tb_cap <- grid.arrange(rmg, table, cap,
#                            heights=c(1.1, 0.3, 0.3), widths=c(0.2, 2),
#                            layout_matrix=rbind(c(1, 1), c(NA, 2), c(3, 3)))
# 
# ggsave('1_base_model_totals_caption.jpg',
#        plot=rmg_tb_cap, height=6.5, width=6, unit ='in')
```
